---
title: "Assignment 5A: Airline Delays"
author: "Mubin Ejaz"
format: html
editor: visual
---

## CodeBase

I had decided to read in the image and extract data to continue but ended up wasting so much time.

At the end I did have to take ChatGPT's advice and just save the data in a csv and read it. That way it was much cleaner to read, plot, analyse, and conclude.

## Code Chunk I tried

```{r loading libraries}
library(knitr)
library(dplyr)
library(tidyverse)
library(tesseract)

```

```{r}
df2<-read_csv("5A.csv")

airline_tidy <- df2 |>
  pivot_longer(
    cols = `Los Angeles`:Seattle,
    names_to = "Destination",
    values_to = "Flights"
  )
airline_tidy

# Using pivot_longer seemed easier to manipulate the data. 


```

```{r}
totals <- airline_tidy |>
  group_by(Airline, Destination) |>
  summarise(
    Total = sum(Flights),
    .groups = "drop"
  )
totals

```

```{r}
delays <- airline_tidy |>
  filter(Status == "delayed")


```

```{r}
delay_rates <- airline_tidy |>
  group_by(Airline, Destination) |>
  summarise(
    Total = sum(Flights),
    Delayed = sum(Flights[Status == "delayed"]),
    Delay_Rate = Delayed / Total,
    .groups = "drop"
  )

```

```{r}
overall <- airline_tidy |>
  group_by(Airline) |>
  summarise(
    Total_Flights = sum(Flights),
    Delayed_Flights = sum(Flights[Status == "delayed"]),
    Delay_Rate = Delayed_Flights / Total_Flights * 100,
    .groups = "drop"
  )
ggplot(delay_rates,
       aes(x = Destination, y = Delay_Rate, fill = Airline)) +
  geom_col(position = "dodge") +
  labs(
    title = "Arrival Delay Rates by Airline and Destination",
    x = "Destination",
    y = "Delay Rate (%)"
  )
```

```{r}
write_csv(overall, "output_5A.csv")
```

Simple data extraction and outputting to FlightDelays.csv gave me a lots of untidy data. All the data has been extracted to one single column and second and third column is just garbage (probably image formatting.

**Analysis:**

I grouped the data by destination and airlines. Calculated the total number of flights. Also calculated the delayed flights. That helped me calculate the delay rate (# of delayed flights/total number of flights). I did need LLM help in doing so, as I had already taken so long and wasted time in trying to extract and manipulate the image.\
\
Conclusion:\
Alaska airline had 501 delay flights, with a delay rate of 13.27. AMWest had 787 delayed flight but the delay rate was much smaller, 10.89.\
I would lean more towards taking Alaska airline.
