---
title: "Global Baseline Estimate"
author: "Mubin"
format: html
editor: visual
---

## Approach

I'll use the data that I collected in last week 2A. The formula provided is:\
\
**Global Baseline Estimate = Mean Movie Rating + Pitch Perfect 2's rating relative to average + Param's** \*\*rating relative to average\*\*\
Predicted Ratings = µ + b~u~ + b~i~

I already have µ as average_rating in my code. I'll calculate Avg User Ratings as\

```{=html}
<pre>

   userrating <- dbGetQuery(
                             MovieRatingConnection,
                             "SELECT users.name, 
                             AVG(ratings.rating) 
                             AS avgUserRatings
                             FROM users 
                             LEFT JOIN ratings 
                             ON users.user_id = ratings.user_id 
                             GROUP BY users.name")
                           </pre>
```

User ratings will help me calculate user bias. Since none of the code is due at this time. My main goal is to 1. Get ratings data into R ==done\
2. Calculate each user's average ===done\
3. Calculate each movie's average ===done\
4. Calculate overall average ==== done\
5. Calculate user adjustments =====done\
6. Calculate movie adjustments ==== I might or might not need it\
7. Use those values to predict a missing rating would be fiddling with R to do this.\
8. Make a movie recommendation ==\>final goal

#### Previous assignment

```{r}

library(DBI)
library(RPostgres)
library(dplyr)

MovieRatingConnection <- dbConnect( RPostgres::Postgres(), dbname = "data607", host = "localhost", port = 5432, user = "postgres", password = Sys.getenv("PG_PASSWORD") )

movieratings <- dbGetQuery(MovieRatingConnection, "
  SELECT
      users.name,
      movies.title,
      ratings.rating
  FROM users
  CROSS JOIN movies
  LEFT JOIN ratings
      ON users.user_id = ratings.user_id
      AND movies.movie_id = ratings.movie_id
")
movieratings

movie_summary<-movieratings |>
  group_by(title) |>
  summarise(
    number_of_ratings = sum(!is.na(rating)),
    average_rating = mean(rating, na.rm = TRUE)
  )

userrating <- dbGetQuery(MovieRatingConnection, "
SELECT
    users.name,
    AVG(ratings.rating) AS avgUserRatings
FROM users
LEFT JOIN ratings
    ON users.user_id = ratings.user_id
GROUP BY users.name")

userrating
movie_summary



OverallAverage <- mean(movieratings$rating, na.rm = TRUE)

result_user<- userrating%>% select(name, avguserratings)%>%mutate(diff=avguserratings-OverallAverage)
result_movie<-movie_summary %>% 
              select(title, average_rating)%>% 
              mutate(moviedifftoavg = average_rating-OverallAverage)
df<-movie_summary%>%select(title, average_rating)

GlobalBaselineEstimate <- OverallAverage +
                                         result_movie %>%
                                         filter(title == "The Housemaid") %>%
                                         pull(moviedifftoavg) +
                                         result_user %>%
                                         filter(name == "Shameer Q") %>%
                                         pull(diff)

GlobalBaselineEstimate


```

I would predict that Shameer Q would rate *The Housemaid* 2.259259 (he's a tough critic).

```{r}
#Pulling all the movies, shameer has NA as
recommendations <- movieratings %>%
  filter(name == "Shameer Q", is.na(rating))


recommendations <- recommendations %>%
  left_join(result_movie %>% select(title, moviedifftoavg),
    by = "title"
  ) %>%
  mutate(
    prediction = OverallAverage +
                 moviedifftoavg +
                 result_user %>%
                   filter(name == "Shameer Q") %>%
                   pull(diff)
  )

recommendations
```

Compare to other movies that Shameer has seen and rated, I predict he'd rate this one higher than others as this movie has been rated highly by the people that have seen it. I predict he'd rate it 2.259 (which is a lot higher than he's rated any movies he's seen).

### Challenge:

Running SQL queries and R code still remains my top challenge, and I believe, is the main goal of this assignment. I am getting better but still takes me long time.\
\
