---
title: "Project 2"
author: Mubin Ejaz
format: html
editor: visual
---

## Approach

I have chosen three datasets for my Project 2. \
1) Average US Tuition from 2004 to 2016, from *tidytuesday.com*\
2) Mortality Rate of certain diseases globally 1990-2016 (from discussion board by Ummay Rukiya)\
3) Life Expectancy of female at birth, (also from world bank, when I looked at the data that Ummay mentioned, I looked at more data that I was interested in)\
4) Life Expectancy of Male at birth (Extra data I would be looking at)\
\
In this approach part of the project, I plan to only read these three datasets as raw data, pivot long and output.

Reading in Average US tuition from 2004 - 2016 and turning wide format to long format. Also outputing to a csv file to see how the semi clean data is looking

```{r}

library(tidyr)
library(tidyverse)

library(dplyr)


# Import raw data

raw_data_avg_tuition <- read.csv("us_avg_tuition.csv")

# Reshape to tidy format

raw_data_avg_tuition <- raw_data |>
  pivot_longer(2:13,
    names_to  = "Years",
    values_to = "income"

  ) |>
  rename_with(tolower) |>
             drop_na()
overall_AvgTuition<- raw_data_avg_tuition |> 
  mutate(years=as.numeric(substr(years,2,5)))

overall_AvgTuition<-overall_AvgTuition |>
                    mutate(income=as.numeric(gsub("[,$]", "", income)))
write.csv(overall_AvgTuition, "CleanedUP_Avg_Tuition.csv")

```

Reading in Mortality Rate and turning it from wide to long format. Also outputing to a csv file to see how the semi clean data is looking

```{r}

raw_data_global_mortality<- read.csv("global_mortalityshareofdeathsbycause2016.csv")

tidy_data2 <- raw_data_global_mortality |>
  
  pivot_longer(4:last_col(),
               
               names_to  = "Disease",
               
               values_to = "Number Of Deaths"
               
  ) |>
  
  rename_with(tolower) |>
  
  drop_na()

write.csv(tidy_data2,"Cleanedup_MortalityRate.csv")

```

Reading in Life Expectancy of women and turning wide format to long. Also outputing to a csv file to see how the semi clean data is looking

```{r}


raw_data_life_expectancy_female<- read.csv("API_LifeExpectancy_F.csv")

tidy_data2 <- raw_data_life_expectancy_female |>
  
  pivot_longer(5:last_col(),
               
               names_to  = "Year",
               
               values_to = "Expected age to die"
               
  ) |>
  
  rename_with(tolower) |>
  
  drop_na()
write.csv(tidy_data2,"CleanedUpLifeExpectancy_Female.csv")



```

Reading in Life Expectancy of men and pivoting long. Also outputing to a csv file to see how the semi clean data is looking

```{r}
raw_data_life_expectancy_male<- read.csv("API_LifeExpectancyM.csv")

tidy_data2 <- raw_data_life_expectancy_male |>
  
  pivot_longer(5:last_col(),
               
               names_to  = "Year",
               
               values_to = "Expected age to die"
               
  ) |>
  
  rename_with(tolower) |>
  
  drop_na()
write.csv(tidy_data2,"CleanedUpLifeExpectancy_Male.csv")

```
