---
title: "Project 1"
author: "Mubin Ejaz"
format: html
editor: visual
---

# Approach:

#### TextFile manupulation:

I will be starting out by saving the .txt provided to the same folder as the project. I'll use the read_lines("textfilename") function of tidyverse/dplyr to read in the text file. I noticed the pattern of the text file to have the following pattern:\
dash row —-\>row 1 (I ended up stripping this one off the bat)\
header —–\>row 1\
header ——\>row 2\
dash row —–\>row 3\
player data—\>row 4\
player data —\>row 5\
dash row—\>row 6\
player data—–\>row 7\
player data——\>row 8\
dash row —–\>row 9\
\
Row 1 to row 3 can now just simply be stripped away by declaring a new variable and start saving it from line 4 of the txt file.

I also can use str_split and use \| as a delimiter to break the rows into columns.\
\
Trying to display playerDataFrame and ratingsDataFram was giving me error. Upon google search, I realize they were still just list of strings and characters the **as.data.frame(do.call(rbind, playerDataFrame))** would help me convert them into actual data frames that I can then manipulate.\
\
I'll have to play around with the code to see which combination of join/ select/mutate would give me a better structure that I can then output using write_csv().

## Code Chunks Tried

```{r}
library(tidyverse)
library(readr)
library(knitr)
library(DT)
```

```{r}
#reading the text file in.  
TextFile <- read_lines("tournamentinfo.txt")

#declaring TF as a character tibble. This is where structured TextFile will be saved
TF <- character()

#counter variable
counter <- 1

#length of TextFile, would help me loop through the raw TextFile
lf <- length(TextFile)

#looping through TextFile from 1 to end row of the TextFile 
while (counter <= lf) {
  #stripping the first row of dashes
  if (!str_detect(TextFile[counter], "^\\s*-+\\s*$")) {
    TF <- c(TF, str_sub(TextFile[counter],2, str_length(TextFile[counter])-1))
  }
  
  counter <- counter + 1
}

r<-1

#will split the user data vs ratings data. 
playerDataFrame<-character()
ratingsDataFram<-character()

lf<-length(TF)
n<-3
i<-1
j<-1

#looping through and splitting rows into different dataframes
while(lf){
  if(n %% 2 == 1){
    playerDataFrame[i]<-TF[n]
    i<-i+1
    
  }
  else if(n%% 2 == 0){
    ratingsDataFram[j]<-TF[n]
    j<-j+1
  }
  n=n+1
  lf=lf-1
}

n<-1
r<-1
while(r<length(playerDataFrame)){
  playerDataFrame[r]<-str_split(playerDataFrame[r], "\\|")
  r<-r+1
}

n<-1
r<-1
while(r<length(playerDataFrame)){
  ratingsDataFram[r]<-str_split(ratingsDataFram[r], "\\|")
  r<-r+1
}


#Trying to display playerDataFrame and ratingsDataFram was giving me error. Upon 
#google search, i realize they were still just list of strings and characters
# the as.data.frame(do.call(rbind, playerDataFrame)) would help me convert them into
#data frames that I can then manipulate

playerDataFrame<-as.data.frame(do.call(rbind, playerDataFrame))
ratingsDataFram<-as.data.frame(do.call(rbind, ratingsDataFram))

ratingsDataFram <- ratingsDataFram %>%
  mutate(PlayerNumber = row_number())


playerDataFrame <- playerDataFrame %>%
  mutate(V1 = as.integer(V1))

combinedData <- playerDataFrame %>%
  left_join(
    ratingsDataFram,
    by = c("V1" = "PlayerNumber")
  )

selectiveData<-combinedData|> select('Name' = V2.x, 'Total Points'=V3.x, 'State' = V1.y, 'federation ID and Average' = V2.y)
 datatable(selectiveData)
```
